---
title: "Chess Tournament"
subtitle: "DATA 607 Project 1"
author: "Jhalak Das"
date: "September 18, 2022"
output: html_document
---

**Project Description**
```
In this project, I was given text a file about a chess tournament result. I have uploaded the file to my github repository and created an .Rmd file with the following information for all of the players:
  
Player’s Name, Player’s State, Total Number of Points, Player’s Pre-Rating, and Average Pre Chess Rating of Opponents

The outcome of the code is in the following format:
Gary Hua, ON, 6.0, 1794, 1605


Here, 1605 was calculated by using the pre-tournament opponents’ ratings of 1436, 1563, 1600, 1610, 1649, 1663, 1716, and dividing by the total number of games played.


All of my codes are in an R markdown file (and published to rpubs.com); with my data accessible for the person running the script.

```

#### Calling Necessary Libraries and Importing Data
```{r}
suppressWarnings(library(readr))
suppressWarnings(library(stringr))
suppressWarnings(library(dplyr))

```

Next I will read in the chess data using readr's read_lines function which loads each line into a character vector.

```{r read_data}
lines <- read_lines("https://raw.githubusercontent.com/JHALAK-sps/Chess/master/tournamentinfo.txt")
lines[1:10]


```


```{r}
theUrl <- 'https://raw.githubusercontent.com/dcorrea614/MSDS/master/tournamentinfo.txt'

dfTournament <- read_csv(file = theUrl,col_names = FALSE)
```


### Interpreting the Data

If we notice, the first 4 lines are not part of the dataset. Followed by it the player information and games played, repeating after every 3 lines. Thus, I, separated the data into 2 matrices to make it a traditional data set.

```{r}
mTournament <- matrix(unlist(dfTournament), byrow=TRUE)

m1 <- mTournament[seq(5,length(mTournament),3)]
head(m1)

m2 <- mTournament[seq(6,length(mTournament),3)]
head(m2)
```


### Regular Expression and String Manupulation
Here I have extracted all the feature values in vector using regular expressions.

```{r Deconstructing}
# matching first numbers
ID <- as.numeric(str_extract(m1, '\\d+'))

# matching the first combination of a letter, any amount of characters and "|"
Name <- str_extract(m1, '[A-z].{1,32}') 

# extracting the name
Name <- str_trim(str_extract(Name, '.+\\s{2,}'))

# matching the first two letters (state) in the second matrix 
State <- str_extract(m2, '[A-Z]{2}') 

# matching at least 1 number, a period, and 1 number
TotalNumberofPoints <- as.numeric(str_extract(m1, '\\d+\\.\\d'))

# matching the combination of "R", any characters and "-"
PreRating <- str_extract(m2, 'R:.{8,}-')

# matching first 4 numbers
PreRating <- as.numeric(str_extract(PreRating, '\\d{1,4}'))

# matching all combinations of 1 letter 2 spaces and any numbers
Rounds <- str_extract_all(m1, '[A-Z]\\s{2,}\\d+')

# matching numbers
Rounds <- str_extract_all(Rounds, '\\d+')
```


### Calculating Average Pre Chess Rating of Opponents

Using two vectors from the previous step, I am able to calculate the Average Pre Chess Rating of Opponents using a loop.

```{r}
AvgOppPreChessRating <- c()
for(i in c(1:length(Rounds))){
  AvgOppPreChessRating[i] <- round(mean(PreRating[as.numeric(Rounds[[i]])]),0)
}
AvgOppPreChessRating

```


## Constructing the Data Frame

```{r}
Project1 <- data.frame(ID,Name,State,TotalNumberofPoints,PreRating,AvgOppPreChessRating)
head(Project1)
```

## Writing the CSV File

```{r}
write_csv(Project1, 'tournament.csv' , append = FALSE)
```










